---
title: "PCGLM"
author: "Lorena Leon"
# output: rmarkdown::html_vignette
output:
  html_document:
    toc: true
    toc_float: true
    # number_sections : true
    theme: simplex
    highlight: tango
vignette: >
  %\VignetteIndexEntry{my-vignette}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---
  
  
```{r include=FALSE}
knitr::opts_chunk$set(
message=FALSE, warning=FALSE, paged.print=TRUE, collapse = TRUE, comment = "#>"
)
```


```{r setup}
library(pack); library(catdata)  # For addiction data
library(mlogit)
```

--------

## Binomial case

--------

## Multinomial case - Non ordered response

### Reference standard case

```{r }

data(addiction)
summary(addiction)
head(addiction)
# vignette("multinomial-addiction1")
addiction$ill<-as.factor(addiction$ill)
data1 <- addiction[,c("ill","gender","university", "age")]
data2 <- na.omit(data1)

summary(data2)

dist1 <- new(ReferenceF)
(mod1 <- dist1$GLMref(
  response = "ill", explanatory_complete = c("intercept"), explanatory_proportional = c("NA"),
  distribution = "logistic", categories_order = c("0", "1", "2"), dataframe = data2
)) # -724.8664

(mod2 <- dist1$GLMref(
  response = "ill", explanatory_complete = c("intercept"), explanatory_proportional = c("NA"),
  distribution = "logistic", categories_order = c("0", "2", "1"), dataframe = data2
)) 
(mod3 <- dist1$GLMref(
  response = "ill", explanatory_complete = c("intercept"), explanatory_proportional = c("NA"),
  distribution = "logistic", categories_order = c("1", "2", "0"), dataframe = data2
)) 
(mod4 <- dist1$GLMref(
  response = "ill", explanatory_complete = c("NA"), explanatory_proportional = c("intercept"),
  distribution = "logistic", categories_order = c("0", "1", "2"), dataframe = data2
)) 
(mod5 <- dist1$GLMref(
  response = "ill", explanatory_complete = c("NA"), explanatory_proportional = c("intercept"),
  distribution = "logistic", categories_order = c("0", "2", "1"), dataframe = data2
))
(mod6 <- dist1$GLMref(
  response = "ill", explanatory_complete = c("NA"), explanatory_proportional = c("intercept"),
  distribution = "logistic", categories_order = c("1", "2", "0"), dataframe = data2
)) 

(mod7 <- dist1$GLMref(
  response = "ill", explanatory_complete = c("gender"), explanatory_proportional = c("NA"),
  distribution = "logistic", categories_order = c("0", "1", "2"), dataframe = data2
)) 
(mod8 <- dist1$GLMref(
  response = "ill", explanatory_complete = c("gender"), explanatory_proportional = c("NA"),
  distribution = "logistic", categories_order = c("0", "2", "1"), dataframe = data2
)) 
(mod8_2 <- dist1$GLMref(
  response = "ill", explanatory_complete = c("gender"), explanatory_proportional = c("NA"),
  distribution = "logistic", categories_order = c("2", "0", "1"), dataframe = data2
)) 
(mod9 <- dist1$GLMref(
  response = "ill", explanatory_complete = c("gender"), explanatory_proportional = c("NA"),
  distribution = "logistic", categories_order = c("1", "2", "0"), dataframe = data2
)) 

dist_10 <- new(ReferenceF)

(mod10 <- dist_10$GLMref(
  response = "ill", explanatory_complete = c("NA"), explanatory_proportional = c("gender"),
  distribution = "logistic", categories_order = c("0", "1", "2"), dataframe = data2
)) 
(mod11 <- dist1$GLMref(
  response = "ill", explanatory_complete = c("NA"), explanatory_proportional = c("intercept", "gender", "university"),
  distribution = "logistic", categories_order = c("0", "1", "2"), dataframe = data2
)) 
(mod12 <- dist1$GLMref(
  response = "ill", explanatory_complete = c("intercept", "gender", "university"), explanatory_proportional = c("NA"),
  distribution = "logistic", categories_order = c("0", "1", "2"), dataframe = data2
)) 
# NO LO DEJA EL OTRO
(mod13 <- dist1$GLMref(
  response = "ill", explanatory_complete = c("intercept", "gender", "university"), explanatory_proportional = c("age"),
  distribution = "logistic", categories_order = c("1", "2", "0"), dataframe = data2
)) 
# FUNCIONA
(mod13a <- dist1$GLMref(
  response = "ill", explanatory_complete = c("intercept", "age", "gender", "university"), explanatory_proportional = c("NA"),
  distribution = "logistic", categories_order = c("1", "2", "0"), dataframe = data2
)) 
(mod13a_1 <- dist1$GLMref(
  response = "ill", explanatory_complete = c("intercept", "age", "gender", "university"), explanatory_proportional = c("NA"),
  distribution = "logistic", categories_order = c("0", "2", "1"), dataframe = data2
)) 
(mod13b <- dist1$GLMref(
  response = "ill", explanatory_complete = c("gender", "university"), explanatory_proportional = c("intercept", "age"),
  distribution = "logistic", categories_order = c("1", "2", "0"), dataframe = data2
)) 

```

Explanatory variables as factors:

```{r}
data2$gender <- as.factor(data2$gender)
data2$university <- as.factor(data2$university)
(mod7 <- dist1$GLMref(
  response = "ill", explanatory_complete = c("gender"), explanatory_proportional = c("NA"),
  distribution = "logistic", categories_order = c("0", "1", "2"), dataframe = data2
))
(mod8 <- dist1$GLMref(
  response = "ill", explanatory_complete = c("gender"), explanatory_proportional = c("NA"),
  distribution = "logistic", categories_order = c("0", "2", "1"), dataframe = data2
))
(mod9 <- dist1$GLMref(
  response = "ill", explanatory_complete = c("gender"), explanatory_proportional = c("NA"),
  distribution = "logistic", categories_order = c("1", "2", "0"), dataframe = data2
))
(mod11 <- dist1$GLMref(
  response = "ill", explanatory_complete = c("NA"), explanatory_proportional = c("intercept", "gender", "university"),
  distribution = "logistic", categories_order = c("0", "1", "2"), dataframe = data2
))
(mod12 <- dist1$GLMref(
  response = "ill", explanatory_complete = c("intercept", "gender", "university"), explanatory_proportional = c("NA"),
  distribution = "logistic", categories_order = c("0", "1", "2"), dataframe = data2
))
```

### Category-specific
#### Travel Mode
##### Example 8.4 Tutz: 
The choice of travel mode of n = 840 passengers in Australia was investigated by Greene (2003). The
data are available from the R package Ecdat. The alternatives of travel mode were air, train, bus, and
car, which have frequencies 0.276, 0.300, 0.142, and 0.280. Air serves as the reference category. As
category-specific variables we consider travel time in vehicle (timevc) and cost, and as the global variable
we consider household income (income). The estimates show that income seems to be
influential for the preference of train and bus over airplane. Moreover, time in vehicle seems to matter for
the preference of the travel mode. Cost turns out to be non-influential if income is in the predictor.

Load data from mlogit library:
```{r}
# library(mlogit)
data(ModeChoice, package = "Ecdat")
head(ModeChoice)
travel.long <- mlogit.data(ModeChoice, choice = "mode", shape = "long", alt.levels = c("air", "train", "bus", "car"))
head(travel.long)
```

With PCGLM
```{r}
head(travel.long)
choice <- sub(".*\\.", "", rownames(travel.long))
indv <- sub("\\..*", "", rownames(travel.long))
travel.long88 <- as.data.frame(cbind(indv, choice, travel.long))
head(travel.long88, 5)

dist7 <- new(ReferenceF)
(exp_8_3 <- dist7$GLMref_ec(
  response = "choice", actual_response = "mode",
  individuals = "indv",
  explanatory_complete = c("intercept", "hinc"),
  depend_y = c("gc", "invt"),
  distribution = "logistic", categories_order = c("train", "bus", "car", "air"), dataframe = travel.long88,
  design = "tutz"
))
```

##### Robustness of Student link function in multinomial choice models
Results from page 8: The log-likelihood obtained with the MNL is −185.91 as obtained by Louviere et al. (2000) page 157.
With PCGLM:
```{r}
dist3 <- new(ReferenceF)
(table3 <- dist3$GLMref_ec(
  response = "choice", actual_response = "mode",
  individuals = "indv",
  explanatory_complete = c("intercept", "hinc", "psize"),
  depend_y = c("gc", "ttme"),
  distribution = "logistic", categories_order = c("air", "train", "bus", "car"), dataframe = travel.long88,
  design = "louviere"
))
(table4 <- dist3$GLMref_ec(
  response = "choice", actual_response = "mode",
  individuals = "indv",
  explanatory_complete = c("intercept"),
  depend_y = c("ttme"),
  distribution = "logistic", categories_order = c("air", "train", "bus", "car"), dataframe = travel.long88,
  design = "louviere"
))
(train_1.35 <- dist3$GLMref_ec(
  response = "choice", actual_response = "mode",
  individuals = "indv",
  explanatory_complete = c("intercept", "hinc", "psize"),
  depend_y = c("gc", "ttme"),
  distribution = "student", categories_order = c("air", "bus", "car", "train"), dataframe = travel.long88,
  design = "louviere"
))
```

#### Heating
Heating is a "horizontal" data.frame with three choice-specific
variables (ic: investment cost, oc: operating cost) and some
individual-specific variables (income, region, rooms)

With PCGLM:
```{r}
data("Heating", package = "Ecdat")
H <- mlogit.data(Heating, shape = "wide", choice = "depvar", varying = c(3:12))
choice <- sub(".*\\.", "", rownames(H))
indv <- sub("\\..*", "", rownames(H))

dat4 <- as.data.frame(cbind(indv, choice, H))
head(dat4, 8)
dist3 <- new(ReferenceF)
(A98 <- dist3$GLMref_ec(
  response = "choice", actual_response = "depvar",
  individuals = "indv",
  explanatory_complete = c("intercept", "income"),
  depend_y = c("oc", "ic"),
  distribution = "logistic", categories_order = c("ec", "er", "gc", "gr", "hp"), dataframe = dat4,
  design = "tutz"
))
```


--------


--------

## Multinomial case - Ordered response

